Effects of resonant and random excitations on the proton beam in the Large Hadron Collider, with applications to the design of pulsed hollow electron lenses for active halo control

### 2018-03-25 resex.tex
Changes to title, abstract, and intro.

\title{Effects of resonant and random excitations on the proton beam
  in the Large Hadron Collider, with applications to the design of a
  pulsed hollow electron lens for active halo control}

@@

\begin{abstract}
  We present the results of numerical simulations and experimental
  studies about the effects of random and resonant excitations on
  proton losses, emittances, and beam distributions in the Large
  Hadron Collider (LHC). In addition to shedding light on complex
  nonlinear effects, these studies are applied to the design of hollow
  electron lenses (HEL) for active beam halo control. In the
  High-Luminosity Large Hadron Collider (HL-LHC), a considerable
  amount of energy will be stored in the beam tails. To control and
  clean the beam halo, the installation of two hollow electron lenses,
  one per beam, is being considered. In standard electron-lens
  operation, a proton bunch sees the same electron current at every
  revolution. Pulsed electron beam operation (i.e., different currents
  for different turns) is also considered, because it can widen the
  range of achievable halo removal rates. For an axially symmetric
  electron beam, only protons in the halo are excited. If a residual
  field is present at the location of the beam core, these particles
  are exposed to time-dependent transverse kicks and to noise. We
  discuss the numerical simulations and the experiments conducted
  in~2016 and~2017 at injection energy in the LHC. The excitation
  patterns were generated by the transverse feedback and damping
  system, which acted as a very flexible source of dipole kicks. The
  main observables were losses, emittances, and transverse
  distributions as a function of excitation patterns and
  strengths. The resonant excitations induced rich dynamical effects
  and nontrivial changes of the beam distributions, which, to our
  knowledge, have not previously been observed and studied in this
  detail. We conclude with a discussion of the tolerable and
  achievable residual fields and proposals for further studies.
\end{abstract}

@@

Considering past, current and future high energy colliders, each new
machine has represented a considerable leap in stored beam energy
(Table~\ref{tab:stored_energy}).

Furthermore, recent measurements at the LHC show that the tails of the
transverse beam distribution are overpopulated compared to a Gaussian
distribution. This results in a considerable amount of energy being
stored in the beam tails. In particular, in the case of the LHC, about
5\% of the beam population is stored in the tails (i.e., above
3.5$\sigma$, where $\sigma$ is the Gaussian rms core beam size),
compared to 0.22\% in an ideal Gaussian distribution, leading to 19~MJ
of stored energy for nominal LHC parameters and 34~MJ in the case of
HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a
mechanism is needed to deplete the beam tails in a controlled
manner. Further information on the needs for halo control in LHC can
be found in Ref.~\cite{helreview}.

@@

In this paper, we concentrate on the experimental and numerical
assessment of possible detrimental effects on the beam core of a
pulsed electron lens. Section~\ref{sec:hel} gives an introduction to
the concept of HELs and summarizes the design parameters of the HL-LHC
HELs. Section~\ref{sec:core} is dedicated to the sources of residual
fields from the HEL in the core region. Sections~\ref{sec:exp}
and~\ref{sec:sim} describe the experimental conditions and the setup
of the numerical tracking simulations. Results and comparisons between
simulations and measurements are given in Section~\ref{sec:simex},
with discussion and summary in Section~\ref{sec:sum}. Auxiliary data
and plots are collected in the Appendices.

>>>>

\title{Effects of resonant and random excitations on the proton beam
  in the Large Hadron Collider, with applications to the design of
  pulsed hollow electron lenses for active halo control}

@@

\begin{abstract}
  We present the results of numerical simulations and experimental
  studies about the effects of resonant and random excitations on
  proton losses, emittances, and beam distributions in the Large
  Hadron Collider (LHC). In addition to shedding light on complex
  nonlinear effects, these studies are applied to the design of hollow
  electron lenses (HEL) for active beam halo control. In the
  High-Luminosity Large Hadron Collider (HL-LHC), a considerable
  amount of energy will be stored in the beam tails. To control and
  clean the beam halo, the installation of two hollow electron lenses,
  one per beam, is being considered. In standard electron-lens
  operation, a proton bunch sees the same electron current at every
  revolution. Pulsed electron beam operation (i.e., different currents
  for different turns) is also considered, because it can widen the
  range of achievable halo removal rates. For an axially symmetric
  electron beam, only protons in the halo are excited. If a residual
  field is present at the location of the beam core, these particles
  are exposed to time-dependent transverse kicks and to noise. We
  discuss the numerical simulations and the experiments conducted
  in~2016 and~2017 at injection energy in the LHC. The excitation
  patterns were generated by the transverse feedback and damping
  system, which acted as a flexible source of dipole kicks. Proton
  beam losses, emittances, and transverse distributions were recorded
  as a function of excitation patterns and strengths. The resonant
  excitations induced rich dynamical effects and nontrivial changes of
  the beam distributions, which, to our knowledge, have not previously
  been observed and studied in this detail. We conclude with a
  discussion of the tolerable and achievable residual fields and
  proposals for further studies.
\end{abstract}

@@

In circular accelerators and storage rings, beam quality can be
affected by the interplay of external excitations with machine
lattice. This work, through calculations and experiments, focuses on
how a certain class of resonant excitations influences beam dynamics
and to which extent these excitations can cause beam losses, emittance
growth, or changes in the particle beam distributions. Besides their
general relevance to the topic of complex nonlinear dynamics, these
studies were motivated by the need to assess the effects of a pulsed
hollow electron lens for active beam halo control.

Considering past, current and future high energy colliders, each new
machine has represented a considerable leap in stored beam energy
(Table~\ref{tab:stored_energy}). Furthermore, recent measurements at
the LHC show that the tails of the transverse beam distribution are
overpopulated compared to a Gaussian distribution. This results in a
considerable amount of energy being stored in the beam tails. In
particular, in the case of the LHC, about 5\% of the beam population
is stored in the tails (i.e., above 3.5$\sigma$, where $\sigma$ is the
Gaussian rms core beam size), compared to 0.22\% in an ideal Gaussian
distribution, leading to 19~MJ of stored energy for nominal LHC
parameters and 34~MJ in the case of
HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a
mechanism is needed to deplete the beam tails in a controlled
manner. Further information on the needs for halo control in LHC can
be found in Ref.~\cite{helreview}.

@@

In this paper, we concentrate on the experimental and numerical
assessment of possible detrimental effects on the beam core of a
pulsed electron lens. Section~\ref{sec:hel} gives an introduction to
the concept of HELs and summarizes the design parameters of the HL-LHC
HELs. Section~\ref{sec:core} is dedicated to the sources of residual
fields from the HEL in the core region. Sections~\ref{sec:exp}
and~\ref{sec:sim} describe the experimental conditions and the setup
of the numerical tracking simulations. Results and comparisons between
simulations and measurements are given in Section~\ref{sec:simex},
with discussion and summary in Section~\ref{sec:sum}.
